### 2018-05-08 Thesis Manuscript/manuscript.tex

\documentclass[man]{apa6}

\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}

% Table formatting
\usepackage{longtable, booktabs}
\usepackage{lscape}
% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
\usepackage{multirow}		% Table styling
\usepackage{tabularx}		% Control Column width
\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
\usepackage{threeparttablex}            % Lets threeparttable work with longtable

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\begin{center}\begin{threeparttable}}
%   {\end{threeparttable}\end{center}\end{landscape}}

\newenvironment{lltable}
  {\begin{landscape}\begin{center}\begin{ThreePartTable}}
  {\end{ThreePartTable}\end{center}\end{landscape}}

  \usepackage{ifthen} % Only add declarations when endfloat package is loaded
  \ifthenelse{\equal{\string man}{\string man}}{%
   \DeclareDelayedFloatFlavor{ThreePartTable}{table} % Make endfloat play with longtable
   % \DeclareDelayedFloatFlavor{ltable}{table} % Make endfloat play with lscape
   \DeclareDelayedFloatFlavor{lltable}{table} % Make endfloat play with lscape & longtable
  }{}%

% The following enables adjusting longtable caption width to table width
% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand\getlongtablewidth{%
 \begingroup
  \ifcsname LT@\roman{LT@tables}\endcsname
  \global\longtablewidth=0pt
  \renewcommand\LT@entry[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}%
  \@nameuse{LT@\roman{LT@tables}}%
  \fi
\endgroup}

\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            pdfauthor={},
            pdftitle={Personalised Map Matched Imputation: imputing missing data from smartphone location logs},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=black,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

\setlength{\parindent}{0pt}
%\setlength{\parskip}{0pt plus 0pt minus 0pt}

\setlength{\emergencystretch}{3em}  % prevent overfull lines

% Manuscript styling
\captionsetup{font=singlespacing,justification=justified}
\usepackage{csquotes}
\usepackage{upgreek}

 % Line numbering
  \usepackage{lineno}
  \linenumbers

\usepackage{tikz} % Variable definition to generate author note

% fix for \tightlist problem in pandoc 1.14
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% Essential manuscript parts
  \title{Personalised Map Matched Imputation: imputing missing data from
smartphone location logs}

  \shorttitle{Missing Data in Smartphone Location Logs}

  \author{Boaz Sobrado\textsuperscript{1}}

  % \def\affdep{{""}}%
  % \def\affcity{{""}}%

  \affiliation{
    \vspace{0.5cm}
          \textsuperscript{1} Utrecht University  }

  \authornote{
    Department of Methodology \& Statistics

    Submitted as a thesis manuscript conforming to APA manuscript guidelines
    (6th edition).

    Correspondence concerning this article should be addressed to Boaz
    Sobrado, . E-mail:
    \href{mailto:boaz@boazsobrado.com}{\nolinkurl{boaz@boazsobrado.com}}
  }

  \abstract{Personal mobility, or how people move in their environment, is
associated with a vast range of behavioural traits and outcomes, such as
socioeconomic status, personality and health. The widespread adoption of
location-sensor equipped smartphones has generated a wealth of objective
personal mobility data. Nonetheless, smartphone collected personal
mobility data has remained underused in behavioural research, partly due
to the practical difficulties associated with obtaining the data and
partly because of the methodological complexity associated with
analysing it. Recent changes in European regulation have made it easier
for researchers to obtain this data, but the methodological difficulties
remain. The difficulty lies in that smartphone location data is
irregularly sampled, sparse and often inaccurate. This results in a high
proportion of missing data. In this paper we present a method called
Personal Map Matched Imputation (PPMI) to deal with missing data and
noise in smartphone location logs. The main innovation of PPMI is that
it creates a personalised spatial map for each individual based on all
the available data. In doing so PPMI leverages the regularity of human
mobility in order to smoothen noisy measurements and impute missing data
values. By simulating missing periods in real data we find that a simple
implementation of PPMI performs as well as existing methods for short (5
minute) missing intervals and substantially better for longer (1 day)
missing intervals. When imputing a subset of real missing data where
travel logs are available as a reference points, we find that PPMI
performs substantially better than existing models.}
  \keywords{Missing Data, Measurement Bias, GPS, Human Mobility \\

    \indent Word count: 8319
  }

\usepackage{amsthm}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}{Lemma}[section]
\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]
\newtheorem{corollary}{Corollary}[section]
\newtheorem{proposition}{Proposition}[section]
\theoremstyle{definition}
\newtheorem{example}{Example}[section]
\theoremstyle{definition}
\newtheorem{exercise}{Exercise}[section]
\theoremstyle{remark}
\newtheorem*{remark}{Remark}
\newtheorem*{solution}{Solution}
\begin{document}

\maketitle

\setcounter{secnumdepth}{0}

\section{Introduction}\label{introduction}

Why is human mobility important? Human mobility affects a wide range of
outcomes, such as health, income and social capital (Goodchild \&
Janelle, 2010). Human mobility measures are quantified metrics of how
people move about in their environment. The most widely administered
personality questionnaires ask individuals to what extent they agree
with statements such as \enquote{I love large parties}, \enquote{I
prefer going to the movies to watching videos at home} and \enquote{I
love to travel to places that I have never been before} (Goldberg et
al., 2006). At their root these questions are mobility measures. In
economic research, the postal code of an individual's home address is
often used as a proxy for socioeconomic status (e.g. Villanueva \&
Aggarwal, 2013). Economists are interested not only in where people
live, but also where they go to work: geographic labour mobility is
related to the income of individuals, the well-being of a country's
economy and even informs the policy of bodies such as the European
Comission (Tatsiramos, 2009). The extensive use of measures like these
across different domains in social science strongly suggests that
mobility metrics are linked to important real world outcomes.
Behavioural researchers have found that mobility measures can be used to
predict academic performance (R. Wang, Harari, Hao, Zhou, \& Campbell,
2015), the incidence of obesity (Zenk, Schulz, \& Odoms-Young, 2009) or
even the onset of a depressive episode in bipolar depression patients
(Palmius et al., 2017). To fully understand behaviour not only must we
understand how behaviour can vary within and between individuals, but
also across environments.

Despite the importance of mobility measures, the majority of metrics are
obtained through the use of questionnaires (such as the aforementioned
personality questionnaire) or specifically through pen-and-paper travel
diaries. As a thought experiment we encourage the reader to try to
remember where exactly he or she was five weeks ago on Saturday at
midday. Most people find this sort of task difficult (if not
impossible), which illustrates the main methodological flaws of
questionnaires and travel diaries. Travel diaries are burdensome to
collect because participants must be explicitly asked to write down on
their movement patterns at frequent intervals (lest they forget them).
The burden to the participants makes data collection expensive, which in
turn limits the duration of data collection in practice. In addition,
the frequent reporting duties of the participant may bias the
participant's behaviour. The limits of human cognition also limit the
accuracy of self-reported measures. There is evidence that statistics
derived from self-reported mobility measures are systematically biased.
For instance, participants under-report the frequency of short trips
(Wolf, Oliveira, \& Thompson, 2003) and underestimate the duration of
regular commutes (Delclòs-Alió, Marquet, \& Miralles-Guasch, 2017).
These obstacles can only be overcome by using objective data on human
mobility.

Social scientists now have the unprecedented opportunity to easily
obtain objective data on human mobility from smartphones. Before
smartphones the only way to collect objective data on human mobility
involved giving participants an expensive professional-grade location
sensor and convincing them to take it with themselves at all times.
Barnett and Onnela (2016) points out that introducing a new device to
the participant's life may bias their behaviour. Moreover, collecting
data in such a way is costly, places a high burden on participants and
therefore the logs do not span a long time (Barnett \& Onnela, 2016).
Today millions of individuals carry smartphones with themselves every
day and do not need to be encouraged to do so by researchers.
Smartphones are equipped with a range of sensors that can be used to
track the location of the device at all times. These smartphones can
collect and store hundreds of location measurements a day. For instance,
Google Location History contains movement information on millions of
users, often spanning years (\emph{Timeline}, 2017). Moreover, recent
changes in EU regulations with regard to consumer data-portability
rights ensure that a willing participant should be able to easily share
this information with researchers at no cost to the either the
participant or the researchers (European Commission, 2017). Researchers
now have the ability to easily access the objective movement records of
millions of individuals spanning several years at little cost and
without a significant burden to the participants.

This paper wishes to achieve four objectives: First, we have argued that
understanding human mobility is important. Secondly, we argued that
social scientists should leverage data logs from smartphones to study
human mobility, instead of relying on out-dated pen-and-paper
questionnaires. Now we will explore the practical difficulties in using
smartphone location logs. Finally we will introduce Personal Map Matched
Imputation (PPMI), a method for surmounting these difficulties. We will
compare PPMI to existing methods in the literature.

\section{Background}\label{background}

Smartphone location measures are obtained primarily (but not
exclusively) by Global Positioning System (GPS) measurements. A GPS
sensor uses the distance between a device and several satellites to
determine the location of the device. Although using a GPS sensor is the
most accurate way to establish location on a smartphone, the GPS sensor
is also the most energy consuming sensor on most smartphones (M. Y. Chen
et al., 2006; LaMarca et al., 2005). In order to avoid battery depletion
and to overcome computational constraints smartphones also use
less-accurate heuristics such as WiFi access points and cellphone tower
triangulation. Smartphone location logs contain measurements from all of
these sources, usually in the form of time-stamped latitude, longitude
and accuracy values. The accuracy \(a\) of any given measurement is
given in meters such that it represents the radius of a 67\% confidence
circle (\emph{Timeline}, 2017). In other words, the true location of a
device should be within the radius \(a\) of the measurement 67\% of the
time.

Researchers often develop custom-made tracking applications which
participants are instructed to download on their phone. Alternatively
participants are given a phone to use for a given period of time with
the custom-made tracking app pre-installed. We call location logs
resulting from these custom-made apps \emph{custom logs}. The advantage
of custom logs is that the researchers can adjust tracking parameters,
such as the the frequency of measurements and the sensor used to make
the measurement. The disadvantage of this approach is that researchers
have to develop or adapt a custom-made tracking application (which is
not easy given hundreds of different types of smartphone models),
distribute it among research participants and enforce participation.
Participants may dislike tracking apps because they view them as more
intrusive and these apps regularly drain the battery of the device
(Harari et al., 2016).

We focus on another solution, which is to take advantage of existing
smartphone location logs (\emph{secondary logs}). The advantages are
clear: repositories such as Google's Location History contain years of
information on millions of users (\emph{Timeline}, 2017) and
participants can share the data by clicking of a button. The
disadvantage is that researchers have no control over the tracking
parameters, often resulting in logs with sparse and inaccurate
measurements. Hence, two important challenges in using secondary logs
are how to deal with missing data and measurement noise.

In order to work with secondary logs, researchers need to be able to
handle the data sparsity that leads to missing data. Missing data is a
pervasive issue in secondary logs as it can arise due to several
reasons. Technical reasons include signal loss, battery failure and
device failure. Behavioural reasons include switching the device or its
location services off. As a result, secondary logs often contain wide
temporal gaps with no measurements. For instance, several research
groups studying mental health report missing data rates between 30\% to
50\% (Grünerbl et al., 2015; Palmius et al., 2017; Saeb et al., 2015).
Other researchers report similar trends in different fields (e.g. Harari
et al., 2016; Jankowska, Schipperijn, \& Kerr, 2015). Figure
\ref{fig:longMeasurementsPerDay} shows a secondary log spanning years
where despite its long duration the sparsity of measurements is also
evident.

\begin{figure}
\includegraphics[width=1\linewidth]{img/missin5minExample2018} \caption{Example of missing data over the entire duration of a secondary log. The x-axis denotes time, the y-axis shows how many measurements are made and each point is a five minute window. For this day there were several periods with no information. These points lie on the x-axis.}\label{fig:longMeasurementsPerDay}
\end{figure}

There is no golden standard for dealing with missing data in GPS logs
(Barnett \& Onnela, 2016). Importantly, spatiotemporal data measurements
are auto-correlated in both time and space. This means that common
practices with other types of data, such as mean imputation, are
unsuitable. For example, imagine an individual who splits almost all her
time between work and home. Suppose she spends a small amount of time
commuting between the two along a circular path. Using mean imputation
to estimate her missing coordinates, we impute her to be at the midpoint
between home and work, even though she has never been there. Worryingly,
there is little transparency on how researchers deal with missing data
(Jankowska et al., 2015).

Another methodological problem is related to the noise in the
measurements. The accuracy of smartphone location measurements is
substantially lower than that of professional GPS location trackers
because smartphones often use less accurate sensors. In professional GPS
trackers less than 80\% of measurements fall within 10 meters of the
true location. GPS measures are most inaccurate in dense urban locations
and indoors (S. Duncan et al., 2013; Schipperijn et al., 2014).
Unfortunately for researchers, this is where people in the developed
world spend most of their time. Figure \ref{fig:accuracyPlot} shows how
accuracy can vary as a function of user behaviour, time and location.
Most notably, low accuracy is often (but not always) associated with
movement (see Figure \ref{fig:accuracyPlot2}).

\begin{figure}
\includegraphics[width=1\linewidth]{img/journeyTillMiddayBoaz} \caption{Measurement accuracy of each logged measurement of a morning journey on February 15th 2017. This includes all measurements from midnight to midday. The red circles denote the accuracy of all logged measurement points (the raw data). The points connected in time are connected by a line. The blue line shows the path without the most inaccurate (accuracy > 400 meters) points filtered out. The red line shows the path with all measurements included. In smartphone logs inaccurate location values are interspersed between more accurate location values at higher sample rates per hour. Inaccurate measures are often followed by more accurate measures. There are several recurring low-accuracy points, such as the one in the northwest corner, possibly the result of cellphone tower triangulation.}\label{fig:accuracyPlot}
\end{figure}

\begin{figure}
\includegraphics[width=1\linewidth]{img/accuracyLocShift} \caption{Measures of user activity and measurement accuracy on February 15th 2017.The upper chart shows the distance from the next measured point in meters over the course of the day. The first peak corresponds to the first journey from the user's home to a gym around 8am. The second, smaller peak before 10 reflects a journey from the gym to the nearby lecture theatre. Both journeys can be seen in Figure 2. The large jump between journey 5 and 6 is measurement error. The lower chart shows the accuracy over the course of the day. The figure shows that measurement inaccuracy is sometimes related to the movement of the individual. Stationary accuracy varies depending on phone battery level, wifi connection and user phone use.}\label{fig:accuracyPlot2}
\end{figure}

Noisy data can lead to inaccurate conclusions if it is not accounted
for. Suppose we wish to calculate an individual's movement in a day. A
simple approach would be to calculate the sum of the distances between
each measurement. But if there is noise, the coordinates will vary even
though the individual is not moving. If the measurements are frequent
and noisy, we will calculate a lot of movement, even if the individual
did not move at all! The problem is further complicated because missing
data and noisy measurements are related. Methods used by researchers to
reduce noise, such as throwing out inaccurate measurements (e.g. Palmius
et al., 2017), can exacerbate the severity of the missing data problem.

In this paper we will propose PPMI as a method for dealing with missing
data and measurement error in secondary location logs. PPMI creates a
discrete personalised map based on the individuals mobility history.
Then it assigns each measurement to a discrete location based on the
individuals entire mobility history. Finally, it uses a classification
method to impute missing periods.

\section{Related Work}\label{related-work}

How have researchers dealt with missing data in human mobility logs thus
far? Unfortunately there is no golden standard in how to deal with this
type of missing data. Researchers are generally vague about what
practices they follow (Jankowska et al., 2015). This vagueness is
worrisome as it invites solutions which contain significant researcher
degrees of freedom (Simmons, Nelson, \& Simonsohn, 2011). Most
researchers simply down-sample temporally and remove missing
observations or use some sort of rule-based common sense imputations
(e.g. Palmius et al., 2017). The only principled approach (that we know
of) aiming to solve the issue of missing data in this context
(i.e.~location logs as they relate to human mobility) is that of Barnett
and Onnela (2016). We will explore the methods of Barnett and Onnela
(2016) and Palmius et al. (2017) in detail subsequently, after
introducing other spatiotemporal methods.

A lack in methods for missing data imputation for human mobility
patterns does not imply there is not a vast literature on modelling
movement. The most widespread models are SSMs, therefore we shall detail
a few examples and subsequently argue that they are nonetheless unsuited
for long term human mobility logs. Ecologists have used SSMs to explain
how animals interact with their environment (T. A. Patterson, Thomas,
Wilcox, Ovaskainen, \& Matthiopoulos, 2008). Preisler, Ager, Johnson,
and Kie (2004) uses Markovian movement processes to characterise the
effect of roads, food patches and streams on cyclical elk movements. The
most well studied SSM is the Kalman filter, which is the optimal
algorithm for inferring linear Gaussian systems. The extended Kalman
filter is the de facto standard for GPS navigation (Z. Chen \& Brown,
2013). The advantage of state space models is that they are flexible,
deal with measurement inaccuracy, include information from different
sources and can be used in real time.

For secondary logs the main limitation of SSM implementations is that
they ignore movement routines. For instance, humans tend to go to work
on weekdays and sleep at night. Because SSMs are based on the Markov
property, they cannot incorporate this information. In other words, the
estimated location at a time-point \(t\) is often based only upon
measurements at timepoints \(t\) and \({t-1}\). Hierarchical structuring
and conditioning on a larger context have been suggested as ways to add
periodicity to Markovian models. These solutions are often
computationally intractable or unfeasible (Sadilek \& Krumm, 2016).
Moreover, these models often assume time and space invariance
(i.e.~location is not a direct function of time or space). These
mathematical assumptions are violated in the case of human movement
patterns. For this reason we do not consider existing SSMs to be useful
for imputing missing data in smartphone mobility logs.

In the wider realm of spatiotemporal statistics there are numerous
missing data imputation methods. These often come from climate or
geological research and rely on spatiotemporal auto-correlations. For
instance, Feng, Nowak, O'Neill, and Welsh (2014) estimate missing values
by incorporating similar observed temporal information from the value's
nearest spatial neighbors. The authors illustrate their example using
rainfall data from gauging stations across Australia. Similarly, Z.
Zhang et al. (2017) use a variety of machine learning methods to impute
missing values. The example provided relates to underground water data.
Generally these models assume fixed measurement stations (such as
rainfall gauging stations). For this reason they cannot be easily
applied to missing mobility tracks without significant pre-processing.

On the other hand, a few researchers have explicitly attempted to impute
missing data from human mobility patterns (Barnett \& Onnela, 2016;
Palmius et al., 2017 ; Wu, Fu, Wang, Xiao, \& Fu, 2017). Importantly,
none of them worked with secondary logs, but rather with custom logs
that have higher sampling frequency. Nonetheless we will detail what
they did as informative examples. Palmius et al. (2017) deal with the
measurement inaccuracy in custom logs by removing from the data set all
unique low-accuracy data points with a speed greater than 100 km/h.
Subsequently the researchers down sample the data to a sample rate of 12
per hour using a median filter. Moreover, Palmius et al. (2017) explain:

\begin{quote}
\enquote{If the standard deviation of in both latitude and longitude
within a 1 h epoch was less than 0.01 km, then all samples within the
hour were set to the mean value of the recorded data, otherwise a 5 min
median filter window was applied to the recorded latitude and longitude
in the epoch}.
\end{quote}

Missing data was imputed using the mean of measurements close in time if
the participant was recorded within 500m of either end of a missing
section and the missing section had a length of \(\leq 2h\) or
\(\leq 12h\) if this occurred after 9pm. In cases where the previous
conditions are not met no values are imputed.

Barnett and Onnela (2016) follow a different approach which is, to the
best of our knowledge, the only principled approach to dealing with
missing data in human mobility data. Barnett and Onnela (2016) work with
custom logs where location is measured for 2 minutes and subsequently
not measured for 10 minutes. Barnett and Onnela (2016) handle missing
data by first converting data to mobility traces, which are defined as a
sequence of flights and pauses. Flights are segments of linear movements
and pauses corresponding to periods of time where a person does not
move. Subsequently, the authors impute missing data by:

\begin{quote}
\enquote{simulat{[}ing{]} flights and pauses over the period of
missingness where the direction, duration, and spatial length of each
flight, the fraction of flights versus the fraction of pauses, and the
duration of pauses are sampled from observed data.}
\end{quote}

This method can be extended to imputing the data based on temporally,
spatially or periodically close flights and pauses. In other words, for
a given missing period, the individual's mobility can be estimated based
on measured movements in that area, at that point in time or movements
in the last 24 hours.

It is also worth mentioning that there is a body of work that
incorporates road features and other background information to generate
predictions or imputations. For instance, Liao, Patterson, Fox, and
Kautz (2007) use hierarchical Markov models in combination with
knowledge about the transportation system to bridge the gap between raw
GPS sensor measurements and high level information such as a user's
destination and mode of transportation. This can be then used to impute
or predict missing steps. Along similar lines Wu et al. (2017) use what
they call a Spatial Temporal Semantic Neural Network (STS-NN) to predict
future human movement. While the authors are concerned with prediction
and not imputation, they devised a method called the Spatial Temporal
Semantic (STS) algorithm which converts raw measurements to machine
learning friendly discrete bins. Working with high-frequency
measurements, Wu et al. (2017)'s method down-samples the raw data
temporally and map-matches the resulting bins to discrete points along
pre-established geographical features such as roads and highways. This
minimises measurement error and paves the way for applying machine
learning methods to human mobility problems. Subsequently we will focus
on methods which do not incorporate external information as they are
more generally applicable.

In this section we have argued that there is a lack of established
practices to follow with respect to missing data in human mobility logs.
Moreover, extensively used spatiotemporal methods, such as state space
models (SSMs), are not well suited to deal with human mobility patterns
in secondary logs. Finally we discussed in detail three approaches which
deal explicitly with mobility patterns from custom or secondary logs
(Barnett \& Onnela, 2016; Palmius et al., 2017 ; Wu et al., 2017).

\section{Methodology}\label{methodology}

\subsection{Notation}\label{notation}

Location measurements, such as those produced by GPS sensors, provide us
with coordinates (latitude and longitude) on the surface of the earth,
which is ellipsoid shaped. Projecting three dimensional measurements in
\(\mathbb{R}^3\) onto a two dimensional plane in \(\mathbb{R}^2\)
results in distortion. For clarity, when we use the term distance we
refer to the geodesic distances on an ellipsoid using the WGS84
ellipsoid parameters.

Subsequently let us simplify by assuming that a persons location is on a
two-dimensional Euclidean plane. Let a person's true location on this
two-dimensional plane be \(G(t) = [G_x(t) G_y(t)]\) where \(G_x(t)\) and
\(G_y(t)\) denote the location of the individual at time \(t\) on the
x-axis and y-axis respectively. For simplicity, we can assume that the
x-axis is the longitude and the y-axis is the latitude. Moreover, let
\(D \in \mathbb{R}^2\) be the recorded data containing the latitude and
longitude. In addition, let \(a\) denote the estimated accuracy of the
recorded data. Furthermore, \(G(t)\), \(D\) and \(a\) are indexed by
time labeled by the countable set \(t = t_1 < ... < t_{n+1}\). If
\(D_t = \emptyset\) it is considered \emph{missing}. The measure of
accuracy \(a_t\) is given in meters such that it represents the radius
of a 67\% confidence circle. The reader should keep in mind that a high
accuracy value \(a_t\) for a measurement \(D_t\) means that the
measurement is inaccurate (this can be rather confusing).

\subsection{Personalised Map Matched
Imputation}\label{personalised-map-matched-imputation}

Our algorithm is designed to leverage the periodic nature of human
movement along with the long span of secondary to deal with measurement
sparsity and inaccuracy.

\paragraph{Modelling assumptions}\label{modelling-assumptions}

First, following Barnett and Onnela (2016) we categorise all time-points
\(t\) as either belonging to the set \(P\) (pause) or set \(F\)
(flight). Conceptually pauses can be understood as periods of time where
an individual spends significant amount of continuous time without
moving. Flights are the times where the individual is moving. Let
\(t_a\) be a pause of length \(n\). \[t_a = t_i < ... < t_{i+n}\] Let
\(t_b\) be a pause of length \(m\) such that there is no temporal
overlap between \(t_a\) and \(t_b\) and \(t_b\) is the first pause after
the end of \(t_a\): \[t_b = t_j < ... < t_{j+m}|t_{i+n}< t_j\] Then it
follows that between the two pauses there must be a flight indexed by
\(t_x\) of length \(j-i+n\). \[t_x =  t_{i+n} < ... < t_{j} |t_x \in F\]
We define pause locations \(G(t_a), G(t_b) | t_a, t_b \in P\) as
locations where an individual spends an extended amount of time in the
same space (e.g.~school, home, work, train station, barber shop, bar,
gym). Importantly, our model assumes period and cyclic human movement
such that there are many pauses \(t_{a1},t_{a2},...,t_{an}\) such that
\(G(t_{a1}) = G(t_{a2}) = ... = G(t_{an})\). Moreover,it is possible for
two consecutive pauses \(t_a\) and \(t_b\) to share a location
\(G(t_a) = G(t_b)\) such that \(t_a \ne t_b\). For example, if the
individual leaves home for a run and returns home without stopping
anywhere else.

Let us define as \(Flight^x_{ab}\) the set of all points belonging to a
flight between \(G(t_a)\) and \(G(t_b)\) at time-point \(t_x\).
\[Flight^x_{ab}= G(t_x)|t_x \in F =  \{G(t_{i+n}),...,G(t_j)\}\] Again,
there are many flights \(t_{x1},t_{x2},...,t_{xn}\) such that
\(Flight^{x1}_{ab} = Flight^{x2}_{ab} = ... = Flight^{xn}_{ab}\). Then,
we can define as \(Path_{ab}\) the set of all flights between \(G(t_a)\)
and \(G(t_b)\) at all time-points. For simplicity, we assume that
\(Path_{ab}=Path_{ba}\).

In addition, we consider all measurements \(D(t)\) to be imperfect
measurements of \(G(t)\):

\[ G(t) = D(t) + \text{Measurement Error}  \]

\subsection{Personalised Map Matched Imputation
algorithm}\label{personalised-map-matched-imputation-algorithm}

Our algorithm performs the following steps:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \emph{Map building}: Extract from measurements \(D\) all pause
  locations and path locations to create a personalised map consisting
  of discrete bins.
\item
  \emph{Binning}: Assign each measurement \(D\) to a unique discrete
  location.
\item
  \emph{Imputing}: Use a classification method to predict missing
  measurements based on all the available information.
\end{enumerate}

\subsubsection{Map building}\label{map-building}

Following Wu et al. (2017)'s spatial-temporal-semantic (STS) feature
extraction algorithm our aim is to transform pause and path locations
into machine learning friendly discrete locations, which we will call
location bins. There are multiple ways of extracting such measurement
clusters in the literature, such as Spatio-Temporal Density-Based
Spatial Clustering of Applications with Noise (ST-DBSCAN) and sequence
oriented clustering (SOC) (Birant \& Kut, 2007). We will focus on
methods which explicitly deal with mobility patterns extracted from
smartphone logs (Barnett \& Onnela, 2016; Palmius et al., 2017 ). Both
of these methods pre-process the data and subsequently use two steps to
extract pause locations: first they extract pauses and their
corresponding locations, then they cluster pause locations based on
spatial proximity. This implementation of PMMI uses a stricter version
of Barnett and Onnela (2016)'s approach to extract pauses.

First the measurements \(D\) are filtered such that only measurements
with an accuracy value lower than \(a_{\text{P lim}}\) remain within the
sample. Then, a measurement \(D_t\) belongs to a pause if and only if:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  The next measurement \(D_{t+1}\) is within \(t_{\text{Pause lim}}\)
  amount of seconds (so it is not missing)
\item
  The next measurement \(D_{t+1}\) is within \(d_{\text{Pause lim}}\)
  meters.
\item
  The duration of the pause is more than \(\delta_{\text{Pause lim}}\)
  seconds.
\item
  Let the measurements of a possible pause which fit the aforementioned
  criteria be \(X = D_{t,t+1,...,t+n}\). Let \(\bar{x}\) be the mean
  coordinates of \(X\), let \(x_{\text{dist}} \in X\) be the furthest
  away points from \(\bar{x}\) and \(\bar{a}\) be the the mean accuracy
  of \(X\). Then \(X\) is a pause if and only if the distances between
  \(\bar{x}\) and \(x_{\text{dist}}\) are \(\leq 2 \times \bar{a}\).
\end{enumerate}

The resulting set of points were then hierarchically clustered using a
distance matrix, such that all points within \(d\) meters of each other
were clustered into a pause location. Each pause location is a bin.

For all remaining measurements we assume that they belong to paths. In
this implementation of PMMI we use the following algorithm to estimate
paths:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Take all measurements which are not pauses, filter them based on an
  accuracy threshold \(a_{\text{Path Lim}}\).
\item
  Create a distance matrix for all remaining measurements \(D_t \in F\)
  and hierarchically cluster it accordingly, such that all points within
  \(d_{\text{Path Lim}}\) meters of each other are clustered into a
  single flight bin.
\end{enumerate}

At this point all empirically observed path bins and pause bins are
extracted. However, there may be some overlap between pause bins and
path bins. Thus, the bins are clustered again, such that the pause bins
retain priority. This means that if a pause bin and a path bin are
within less than \(d\) meters of each other, the path bin is removed.
The reasoning for this is that the threshold for not being in a pause
bin should be higher, as individuals spend the majority of time at a
pause bin. The end result is a discretised map which contains pause and
flight bins based on the entire log history of the individual.

\subsubsection{Binning}\label{binning}

Wu et al. (2017)'s spatial-temporal-semantic (STS) feature extraction
algorithm uses map matching as a ground truth to assign noisy
measurements into discrete bins along roads. In other words, in addition
to the measured data they also use a geographic database that contains
information about the area in which the individual is (e.g.~where
precisely the roads are), and sort measurements into bins based on both
the measurement and the geographic data base. For example, if an
individual is measured as moving closely in parallel to a road A in an
area where there is no other parallel road, Wu et al. (2017)'s method
will assume that the individual is on the road A.

PMMI uses a similar logic, but without using any external geographic
database. The key modification in PMMI is that whilst Wu et al. (2017)
uses an external map, we use the total location history of the
individual to create a personalised map. This map is subsequently used
to bin measurements. This is feasible for two reasons: humans tend to
have repetitive movement habits and secondary logs tend to be long. To
put it in simpler terms, we consider each measurement at \(D_x\) as a
sample of \(Path_{ab}\), and by aggregating many measurements we can use
them to map out \(Path_{ab}\).

Thus, all measurements \(D\) are then assigned to a discrete bin on the
personal map. This includes previous measurements which were discarded
from the map building exercise due to an accuracy \(a\) value which
exceeded \(a_{\text{P lim}}\) or \(a_{\text{F lim}}\). In this
implementation we used a simple assigning function, whereby the
measurements where assigned to the nearest bin.

\subsubsection{Classification}\label{classification}

At this point the objective of PMMI is to take all the information
available about the mobility history of an individual and impute the
missing value. In this implementation we trained an artificial neural
network (ANN) to do so. The ANN has a simple architecture: it consists
of an input layer, a hidden layer and an output layer. The input layer
and the hidden layer employ a rectified linear unit activation function
and the output layer uses a softmax activation function.

The input variables to the ANN are:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  The previous and subsequent observed bin as a binary class matrix.
\item
  The distance in time to the next \& previous bin.
\item
  The time of the day encoded as a cyclical two-dimensional feature.
\item
  The day of the week as a binary class matrix.
\item
  The month of the year as a binary class matrix.
\end{enumerate}

For the encoding of the time of day we took the cosine and the sine
transforms of the amount of seconds that have elapsed after midnight
(London, 2016). This is necessary for the model to be able to understand
that one second past midnight and one second before midnight are in fact
two seconds away from each other. Moreover we scaled the non-binary
values to occupy a range between 0 and 1 in order to ensure convergence.

For a missing time-point at \(D_t \in \emptyset\), the output of the
model is a set of probability estimates associated with every location
bin. That is, for each missing time-point the model returns a vector of
probability estimates (with one estimate per bin) associated with where
the individual is.

\subsection{Datasets \& Analyses}\label{datasets-analyses}

The secondary location log used to train the imputation methods was
collected between 2013 and 2017 on different Android devices from a
single individual. About 54\% of the data is missing for the entire
duration of the log. There are several long periods with no measurements
whatsoever, as well as periods with far less missing data. For days
which were not entirely missing, approximately 22\% of all five minute
segments were missing. The structure of missingness of a non-missing day
is shown in Figure \ref{fig:measurementsPerDay}. As you can see, there
are several long periods over the course of the log for which there are
no measurements. The median sampling frequency per day for non-missing
days is around 0.006 Hz.

\begin{figure}
\includegraphics[width=1\linewidth]{img/missingdayBoaz5min} \caption{Missing data for the entire duration of the log. The x-axis denotes time, the y-axis shows how many measurements are made and each point is a five minute window. The entire log contains several long periods with no information. These points are filled with red and lie on the x-axis.}\label{fig:measurementsPerDay}
\end{figure}

For simplicity, we subsequently used a time period when the individual
was living in the Netherlands. This subset contains 156,000 measurements
over a period of less than six months.

Palmius et al. (2017)'s algorithm was implemented in R based on the
original MATLAB code and pseudocode kindly provided by the authors.
Barnett and Onnela (2016)'s method was slightly adapted in R to fit the
data structure from the original R code provided by the authors. When
executing their models we used the same parameters as the authors did.
To represent Barnett and Onnela (2016)'s model we used the variant where
movements were sampled based on spatial proximity.

All analyses were performed using R (Version 3.4.3; R Core Team,
2017)\footnote{We, furthermore, used the R-packages \emph{bindrcpp}
  (Version 0.2; Müller, 2017), \emph{dbplyr} (Version 1.2.0; Wickham \&
  Ruiz, 2018), \emph{dplyr} (Version 0.7.4; Wickham, Francois, Henry, \&
  Müller, 2017), \emph{geosphere} (Version 1.5.7; Hijmans, 2017a),
  \emph{ggplot2} (Version 2.2.1; Wickham, 2009), \emph{ggthemes}
  (Version 3.4.0; Arnold, 2017), \emph{kableExtra} (Version 0.7.0; Zhu,
  2018), \emph{keras} (Version 2.1.4; Allaire \& Chollet, 2018),
  \emph{knitr} (Version 1.20; Xie, 2015), \emph{leaflet} (Version 2.0.0;
  Cheng, Karambelkar, \& Xie, n.d.), \emph{padr} (Version 0.4.0; Thoen,
  2017), \emph{papaja} (Version 0.1.0.9709; Aust \& Barth, 2018),
  \emph{raster} (Version 2.6.7; Hijmans, 2017b), \emph{RColorBrewer}
  (Version 1.1.2; Neuwirth, 2014), \emph{readr} (Version 1.1.1; Wickham,
  Hester, \& Francois, 2017), \emph{rgdal} (Version 1.2.16; R. Bivand,
  Keitt, \& Rowlingson, 2017), \emph{scales} (Version 0.5.0; Wickham,
  2017), \emph{sp} (Version 1.2.7; Hijmans, 2017a; Pebesma \& Bivand,
  2005), \emph{tibbletime} (Version 0.1.0; Vaughan \& Dancho, 2018), and
  \emph{tidyr} (Version 0.8.0; Wickham \& Henry, 2018).}.

All the code is available on a public repository\footnote{\href{https://github.com/sobradob/thesis}{Github}
  (Sobrado, 2018)} .

\subsection{Results \& Evaluation
Metrics}\label{results-evaluation-metrics}

The results will consist of multiple steps:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Evaluating the performance of the map building and assigning
  functions.
\item
  Comparing the performance of PMMI compared to alternative methods
  (Barnett \& Onnela, 2016; Palmius et al., 2017) and a baseline model
  using cross-validation.
\item
  Comparing the performance of PPMI compared to the aformentioned
  methods using objective ground-truth data (public transportation
  time-stamped locations).
\end{enumerate}

\subsubsection{Map building \& binning
evaluation}\label{map-building-binning-evaluation}

Before we can evaluate the accuracy of the imputations, it is essential
to evaluate how well noise in the data has been smoothed. All methods
smoothen noise differently in an attempt to extract true location
\(G(t)\) from measurements \(D_t\). In the subsequent cross-validation
step we will remove random smoothened 5 minute periods and compare the
imputed value to the smoothened value. For this reason we want to ensure
that the smoothened value actually does reflect an approximation to the
true location \(G(t)\). Otherwise we run the risk of over-estimating the
accuracy of a model in the sense that we will compute the extent to
which an imputation method can correctly impute a smoothened value which
has little to do with the true location \(G(t)\).

In order to evaluate the map building and binning we will first visually
evaluate the paths and pause locations. A visual evaluation of paths
superimposed on a map is an established way to heuristically check their
accuracy (e.g. Brunsdon, 2007). Then, let the average distance between
the actual measured point and the binned point be the \emph{deviation
distance} \(\delta_{\text{dev}}\). With respect to the deviation
distance \(\delta_{\text{dev}}\), we expect:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  A positive relationship between the deviation distance and the
  accuracy \(a\) (in meters) of each measurement.
\item
  Roughly 67\% of the deviation distances \(\delta_{\text{dev}}\) are
  within accuracy \(a\) of each measurement.
\end{enumerate}

\subsubsection{Imputation algorithm performance in
cross-validation.}\label{imputation-algorithm-performance-in-cross-validation.}

We will compare the performance of PMMI, Palmius et al. (2017) and
Barnett and Onnela (2016). In addition, we will also compute two
baseline models. The first one is a naive model, which simply imputes
the previous observed value. The second one we will call the
\enquote{home} model, which imputes that the individual is at home at
all missing timepoints. Using these two models as baselines makes sense
given the high degree of autocorrelation in the data and that
individuals tend to spend most of their time at home. To compare the
performance of the aforementioned methods we will remove 25\% of
measured time intervals at random within a four week period. We will
make our comparisons for intervals of 5 minutes, 1 hour and 1 day. In
other words, we will remove 25\% of time intervals at random, while
varying the duration of the time intervals removed.

For the Barnett and Onnela (2016) and PMMI models we will use all the
available data to train the models with the exception of the time
periods being investigated. Palmius et al. (2017)'s model does not
require training.

To compare all methods with each other we will compute a distance
measure (how far was the removed location from the predicted location)
and the coverage. The coverage refers to the percentage of missing cases
which were imputed. For PMMI's imputations we will use a weighted mean
for the distance measures whereby each 5 minute period is weighted
equally. This is necessary because the other two models downsample
temporally to 5 minutes and because the frequency of measurements in a 5
minute period is correlated with imprecise measurements and visiting
infrequently travelled locations. Hence weights must be used to avoid
inflating error rates for PPMI in comparison to the other models.

In addition, other measures of interest for PMMI are \emph{accuracy} (in
what percentage of the cases was the appropriate bin predicted), the
\emph{confidence} and the \emph{distance expectation}. The confidence is
the probability with which the model predicts the most likely bin. For
instance, if the model predicts the missing bin to be bin A with a
probability of 0.9 then we can say this is a high confidence prediction.
Similarly, the distance expectation is the cross-product of the
estimated probabilities that the individual is at any of all given bins
and the distances between the bins to the true bin. For example, if the
true location of an individual is bin A (bin A is 1000 meters away from
bin B) and the model assigns a probability of 0.9 at bin A and 0.1 at
bin B, then the distance expectation would be 100 meters.

\subsubsection{Imputation algorithm performance using external
benchmarks.}\label{imputation-algorithm-performance-using-external-benchmarks.}

We will compare the performance of PMMI, Palmius et al. (2017), Barnett
and Onnela (2016) and the naive baseline model with regards to imputing
real missing measurments. To do so we will use information from the
Dutch public transportation card. The Dutch public transportation
service provides users with time-stamped location data of when and where
they board, change lines or leave public transportation. In this paper
we used 97 such events in a single month.

To be able to make a comparison between models we will remove all
measurements from within the 5 minute period that a time-stamped
measurement is available. Then we will use each model to impute the
location of the individual within that period and compare it to the true
location.

\section{Results}\label{results}

\subsection{Map building \& binning
evaluation}\label{map-building-binning-evaluation-1}

We used the following parameters to extract pauses: an accuracy limit
\(a_{\text{Pause lim}}\) of 250 meters, a time limit
\(t_{\text{Pause lim}}\) of 300 seconds, a distance limit of
\(d_{\text{Pause lim}}\) 50 meters and a minimum pause duration limit
\(\delta_{\text{Pause lim}}\) of 100 seconds. Moreover, to extract path
bins we used the parameters: \(a_{\text{Path Lim}} = 150 \text{meters}\)
and \(d_{\text{Path Lim}} =300 \text{meters}\).

The model was tuned based on a visual inspection of the resulting
personal map meeting the following criteria:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Plausible extraction of pause locations. For example, a pause in a
  field with no path leading to it is less plausible than a pause at a
  train station, airport, or office building.
\item
  A suitable extraction of paths. Ensuring that individual paths are
  logically consistent (e.g.~do not go over rivers or other unplausable
  behaviour).
\item
  Less than a fifth of extracted bins are assigned only one measurement.
  Unique bins are unavoidable as people sometimes visit a location
  briefly and only once, nonetheless they pose problems for the
  classification algorithms.
\end{enumerate}

When selecting parameters there is a trade off is between bias and
precision. This is because an increase in precision (in the form of a
higher resolution of bin locations) comes at the expense of bias
(assigning measurements to bins becomes more difficult). For instance,
by increasing the clustering distance parameter \(d_{\text{Pause lim}}\)
we can extract more valid pause locations at the risk of falsely
categorising certain measurements to the wrong bin. This is illustrated
in Figure \ref{fig:pauseLocUithof1}.

\begin{figure}
\includegraphics[width=1\linewidth]{img/distanceParam} \caption{Example of pause locations in De Uithof university campus using 150 meters (blue) and 400 meters (red) as clustering parameters.}\label{fig:pauseLocUithof1}
\end{figure}

Map building results in a personalised map with pause and path bins. An
excerpt can be seen in Figure \ref{fig:mapClusterMap}. It is important
to remind the reader that PMMI is map agnostic and uses no external
information. Therefore, the close overlap with features on the map, such
as pause bins at relevant buildings and transportation locations, as
well as the flight bins following roads and railway lines indicate a
high degree of precision in personal map building. As expected, PMMI's
path extraction yields greater accuracy for frequently occurring paths.
For example, the frequently travelled Amsterdam-Utrecht railway line has
been extracted almost perfectly, while the less frequently travelled
Utrecht-Enschede line is far sparser.

\begin{figure}
\includegraphics[width=1\linewidth]{img/mapClusterMap} \caption{Excerpt of the personalised binned map of an individual. Red points are pause locations, blue points are path locations.}\label{fig:mapClusterMap}
\end{figure}

For the entire period examined period the we find a mean deviance
\(\delta_{\text{dev}}\) of 40 meters and a median deviance of 15 meters.
Around 69\% of the deviance values are within their corresponding
accuracy value, which is close to the expected theoretical 67\%.
Approximately 9\% of values were not taken into account when creating
the bins, given that their the accuracy \(a\) exceeded
\(a_{\text{Path Lim}} = 300\).

In comparison, Palmius et al. (2017)'s method has a median deviance
\(\delta_{\text{dev}}\) of 3 meters, with a mean of 115 due to high
deviance outliers. On the other hand, the Barnett and Onnela (2016)
method has mean deviance \(\delta_{\text{dev}}\) of 343 meters and a
median deviance of 8 meters. Barnett and Onnela (2016)'s deviance is
necessarily higher than Palmius et al. (2017), as they down-sample
temporally (like Palmius et al., 2017) and subsequently aggregate into
pauses and linear flights.

The key difference between temporal and spatial down-sampling is shown
in Figure \ref{fig:palmiusVsMe1}. Temporal down-sampling is much more
sensitive to noise in sparsely measured periods because it averages out
values within five minute periods. Often there are only a few noisy
measurements in those periods (see Figure
\ref{fig:longMeasurementsPerDay} which shows that a large proportion of
5 minute periods have less than 3 measurements), which leads to a noise
in the down-sampled values. Unsurprisingly, there is a positive
relationship between deviance and the amount of measurements in each
down-sampled interval. The fact that over 90\% of deviance values are
within accuracy (substantially higher than the expected theoretical
67\%) confirms that temporal down-sampling is not sufficiently filtering
out the noise.

\begin{figure}
\includegraphics[width=1\linewidth]{img/tempVsspace} \caption{The difference between temporal and spatial downsampling. The blue circles are raw measurments, the yellow circles are temporally downsampled locations. Spatially downsampled locations are in red. Due to measurement sparcity and inaccuracy many of the temporally downsampled locations are in unfeasable locations.}\label{fig:palmiusVsMe1}
\end{figure}

\subsubsection{Imputation algorithm performance evaluation in
cross-validation.}\label{imputation-algorithm-performance-evaluation-in-cross-validation.}

The first factor we must take into account is the coverage of each
imputation model. The results show us that while PPMI and the baseline
models impute all missing values, the other two models have difficulties
doing so (Table 1). In particular, the models of Palmius et al. (2017)
and Barnett and Onnela (2016)'s fail to impute an increasing amount of
missing values with an increase in the duration of the missing period.
Palmius et al. (2017)'s method in particular failed to impute a single
missing time period when the duration of the removed periods was one
day. The longer a log, the higher the probability that long time periods
will be missing. Secondary logs are typically long. Hence missing data
imputation methods designed to work with secondary logs must be capable
of imputing longer time periods.

\begin{lltable}
\begin{longtable}{llllllllll}\noalign{\getlongtablewidth\global\LTcapwidth=\longtablewidth}
\caption{\label{tab:resulttable}Distance in meters between the removed time period and the imputed value.}\\
\toprule
 & \multicolumn{3}{c}{Five minutes} & \multicolumn{3}{c}{One Hour} & \multicolumn{3}{c}{One Day} \\
\cmidrule(r){2-4} \cmidrule(r){5-7} \cmidrule(r){8-10}
 & \multicolumn{1}{c}{Mean} & \multicolumn{1}{c}{Median} & \multicolumn{1}{c}{Coverage} & \multicolumn{1}{c}{Mean} & \multicolumn{1}{c}{Median} & \multicolumn{1}{c}{Coverage} & \multicolumn{1}{c}{Mean} & \multicolumn{1}{c}{Median} & \multicolumn{1}{c}{Coverage}\\
\midrule
Barnett \& Onella & 240 & 5 & 97\% & 33 & 6 & 88\% & 62 & 6 & 85\%\\
Palmius & 43 & 9 & 97\% & 497 & 4 & 89\% & NA & NA & 0\%\\
PPMI & 269 & 0 & 100\% & 908 & 0 & 100\% & 5,757 & 0 & 100\%\\
Naive Baseline & 426 & 0 & 100\% & 1,502 & 0 & 100\% & 14,266 & 1,288 & 100\%\\
Home Baseline & 5,599 & 0 & 100\% & 5,667 & 0 & 100\% & 5,757 & 0 & 100\%\\
\bottomrule
\end{longtable}
\end{lltable}

With respect to the distance metrics, Table 1 shows the results of the
distance metrics for each method. We can see that PPMI predicted the
location perfectly for at least about half of the removed values in all
three time categories. To be precise PPMI's prediction accuracy was
88\%, 73\% and 47\% for the 5 minute, 1 hour and 1 day periods
respectively. As a comparison, the baseline naive model's prediction
accuracy was 87\%, 68\% and 22\%, and the home baseline model's were
46\%, 52\% and 40\%. Although the home baseline model and PPMI achieved
similar distance metrics in the 1 day condition, this does not mean that
PPMI defaulted to simply predicting the individual is home at all times
(only 40\% of predictions were home). PPMI outperformed the alternative
models when taking into account the median scores, but not the mean. The
explanation is that the alternative models failed to impute more
difficult time periods. This increases the magnitude of errors of the
PPMI method.

On the other hand, the confidence score for each time period were
heavily skewed left, with a median of 1. Accordingly, the distance
expectation (the crossproduct of the model probabilities of an estimate
with the distances of the true value to the estimates) values were quite
similar to the distance scores, albeit approximately 5\% higher. Based
on this we can say that this implementation of PPMI was overconfident in
its predictions.

\subsubsection{Imputation algorithm performance evaluation using
external
benchmarks.}\label{imputation-algorithm-performance-evaluation-using-external-benchmarks.}

Once we removed the measurements within the 5 minute period of each
time-stamp in the public transportation log, we employed each of the
aforementioned methods to impute the location of the individual at the
time-stamped period. Then we calculated the distance metrics between the
true location (based on the public transport log) and the imputed
location. We did not compute the home baseline model as it would have
always been wrong for this subset of data (travelling and being at home
are mutually exclusive).

First, the median distance between the imputed value and the true value
for the baseline naive model was 555 meters (with a mean of 3303). For
this method, the prediction accuracy was 16\%. This means that the
binning method places the individual at the location reported by the
public transportation travel log in the five minutes prior to the
measurement 16\% of the time. Needless to say the coverage was 100\% as
all missing values were imputed.

For the PPMI the median distance metric was 1037 meters (mean 5637). The
prediction accuracy was higher than that of the baseline naive model at
24\%. Although PPMI imputed more points correctly than the naive
baseline model, those it imputed incorrectly were significantly more
incorrect. All missing values were imputed, hence resulting in a
coverage of 100\%.

The same could not be said about Palmius et al. (2017)'s method, as it
failed to impute almost 40\% of the time periods. This results in a
coverage of 60\%. For the time periods the method did impute, there was
a median distance of 1617 meters between the true location and the
imputed value (with a mean of 1517 meters).

Barnett and Onnela (2016)'s method performed better in terms of coverage
at 81\%. The median distance metric was 1342 meters and the mean value
was 5506 meters.

\subsubsection{Example: effect on aggregate
measures}\label{example-effect-on-aggregate-measures}

Social scientists are most interested in aggregating spatiotemporal data
to more socially relevant metrics, such as the amount of time spent at
home, frequency of travels to new locations, the amount of trips made or
distance covered. As an example we calculated the time spent at home of
the user without imputing any missing data and with all three of the
investigated methods. The results can be seen in Figure
\ref{fig:aggrePlot}. Without any form of missing imputation the
individual's time spent at home is unknown for 12\% of the time, which
adds up to almost four days over the course of a 30 day month.

\begin{figure}
\includegraphics[width=1\linewidth]{img/timeUse} \caption{Proportion of time spent at home in March 2017. The raw values are estimated by downsampling temporally the lattitude and longitude for every 5 minute time period in the month. We used each method's own binning method and classified as at home if the downsampled measurement was within 250 meters from home.}\label{fig:aggrePlot}
\end{figure}

The amount of time spent at home has been found to be a reliable
predictor of extroversion (Harari et al., 2016) and the onset of
depressive episodes in bipolar patients (Palmius et al., 2017). The 12\%
of time that remains unaccounted for could mean the difference between
an introvert and an extrovert, or a healthy individual and an individual
suffering of depression. Filling in these gaps is the exactly the sort
of problem that can be solved using missing data imputation.
Interestingly, despite their radically different imputation methods, all
three examined models suggest that the user spent approximately 60\% of
their time at home.

\subsection{Discussion \& Conclusion}\label{discussion-conclusion}

Overall the PMMI performed better than the alternative models,
particularly during longer missing periods and with objective data
gathered from the Dutch public transport service. However, PPMI did not
perform substantially better than the baseline models for all time
periods and with the objective Dutch public transport data, suggesting
that the classification method can be improved. In addition, the
comparison to the performance of the Barnett and Onnela (2016) and
Palmius et al. (2017) models is somewhat unfair, as they were created
for custom logs, not secondary logs. Nonetheless, the comparison remains
valid as they are the closest we found to a missing data imputation
methods in smartphone location logs.

In addition to higher accuracy under the conditions typical of secondary
logs, the advantages of PPMI are increased coverage and flexibility for
missing data imputation, robustness to irregular sampling, the ability
to model complex non-linear interactions in its imputations, and the
ability to use historical records to smooth movement noise.

PPMI's increased coverage and flexibility comes from its ability to make
complex non-linear predictions. For instance, in a given missing period
it might make sense to predict that the individual is either at home, or
at the office, or at a shop with equal probability. While PPMI can make
such an imputation, none of the alternative methods can do this.
Moreover, the ability to take the prediction probability values from the
neural network also helps in dealing with uncertainty. A known-drawback
of single imputation is that it takes an imputed value and treats it as
observed. Simple rule-based methods such as Palmius et al. (2017)'s are
essentially algorithmic single imputation methods. With PPMI it is
possible to model uncertainty using the predicted probabilities of each
estimate. For instance, in the previous example, we could choose to only
take estimates with a high degree of confidence, thus creating
confidence intervals by adding and subtracting the amount of cases where
the location of the individual is ambiguous.

With respect to irregular sampling, alternative methods use temporally
based down-sampling in order to reduce noise. This leads to
deterioration in resolution not only over space, but also over time. A
combination of irregular sampling with the fluctuating accuracy values
can lead to nonsensical results. For instance, consider a case where
there are two inaccurate measurements in movement at 12:00:01 and
12:04:59. Down-sampling over 5 minute periods will lead to a value that
will be the mean of the two inaccurate samples, which is likely to be a
location the individual is certainly not. PPMI instead down-samples
spatially, which ensures that the binned location is one which is
composed of the mean of hundreds of observations, not just the few that
happen within a single period.

While both Barnett and Onnela (2016) and Palmius et al. (2017) use
historical data to smoothen pause locations by clustering pause
locations with a close degree of spatial proximity, neither of them do
the same for non-pause locations. This may be feasible with high
frequency, regularly sampled short duration logs but creates noise with
secondary logs. Moreover, with secondary logs it is feasible to
spatially \enquote{average out} multiple samples of the same path in
order to recreate it in its entirety. For instance, although the mean
sampling frequency during train travels on the Amsterdam-Utrecht line is
low (about 0.01 Hz) the personalised map manages to recreate the train
line almost perfectly, despite being completely map agnostic. Map
agnosticism makes the model more flexible as it can be used for areas
where no good geographical databases are available.

There are multiple methodological limitations in this paper. Most
importantly, the evaluation methods are imperfect. The golden standard
would be to use at least one highly accurate professional grade GPS
device with high sampling frequency to validate results. Until that is
available, the use of public transport data and cross-validation is just
a substitute.

Furthermore, PPMI can be further developed. The map building function,
the assignment function and the classification model remain simplistic
and could be improved.

In map building, the probability of a pause at a given location is
certainly related to other factors, such as the time of the day as well
as the prior history of pauses at that location. These factors are not
taken into account in the pause extraction function. Improved methods
would do well to do so. As for paths, a drawback of the current method
is that the density of the bins is a function of the clustering
parameter \(d\), the distance between the observed points and their
sampling density. It does not take into account the length of the path
as well as the average sampling frequency of the path. This is an issue
because it can lead to bins to which data points are seldom assigned.
For example, while the Amsterdam-Utrecht line has been mapped out almost
perfectly, many of the clusters along the route have only been assigned
few measurements. This leads to difficulties in the classification part
of the model, as infrequently observed clusters are hard for the model
to predict.

The current assignment function simply assigns each measurement to the
nearest bin. It does not take into account any contextual information
that can be gleaned from the entire movement history of the information,
such as what path they are on. For instance, assume that it is known
that an individual is travelling from point A to point B along path AB,
and there is an inaccurate measurement closest to a bin which belongs to
path AC. By only taking distance into account, the measurement can get
assigned to the wrong bin on path AC. An improvement would be use a
Bayesian method, whereby assignment is a function of both the
measurement and a model of the individuals movement history. In terms of
state-space models the state equation would represent a probabilistic
representation of where the individual could be at that given time based
on the individuals entire movement history. The space side of the model
would be a measurement equation representing the measurement and the
uncertainty surrounding it in the form of \(a\). This would be a similar
implementation to that used in Liao et al. (2007), but rather taking the
personalised map instead of information about road networks.

As for the simplicity of the classification method, the neural network
which was used to generate predictions used no information on sequence
patterns longer than the previous and next bin. It also has no
contextual information about how far away the bins are from each other.
A model with more input variables and a more sophisticated design, such
as a recurrent neural network (RNN), or a long short-term memory
recurrent neural network (LSTM) would likely perform significantly
better. Moreover, more inputs can be added from alternative sensors in
smartphones such as the accelerometer, barometer, measures of phone
activity and so on. These can be used to supplement or improve the raw
location measurements, or indeed inform the predictive model.

In conclusion, with this paper we aim to start filling the noticeable
gap in the social science methodology literature in using smartphone
location logs to study human movement. Clearly there is room for
improvement, but this is to be expected with new methods. We are in the
early days of using smartphone location measurements in social science.
Nonetheless, the methodological advantages are clear: millions of
individuals have years long location logs containing objective
measurements. In addition, these measurements can be easily obtained.
This is an unprecedented opportunity. However, with great opportunity
comes a great responsibility, as this data raises difficult questions
about privacy that researchers must be prepared to answer. Privacy
concerns should be addressed, but should not dissuade researchers from
developing new methods. Objective location logs are vastly superior to
alternatives, such as questionnaires and travel diares, which rely on
accurate self-reporting. Social science researchers must take advantage
of regulatory changes with regard to data portability and put the vast
wealth of data collected by commercial entities to scientific use.

\newpage

\section{References}\label{references}

\begingroup
\setlength{\parindent}{-0.5in} \setlength{\leftskip}{0.5in}

\hypertarget{refs}{}
\hypertarget{ref-R-keras}{}
Allaire, J., \& Chollet, F. (2018). \emph{Keras: R interface to
'keras'}. Retrieved from \url{https://CRAN.R-project.org/package=keras}

\hypertarget{ref-R-ggthemes}{}
Arnold, J. B. (2017). \emph{Ggthemes: Extra themes, scales and geoms for
'ggplot2'}. Retrieved from
\url{https://CRAN.R-project.org/package=ggthemes}

\hypertarget{ref-R-papaja}{}
Aust, F., \& Barth, M. (2018). \emph{papaja: Create APA manuscripts with
R Markdown}. Retrieved from \url{https://github.com/crsh/papaja}

\hypertarget{ref-barnett_inferring_2016}{}
Barnett, I., \& Onnela, J.-P. (2016). Inferring mobility measures from
GPS traces with missing data. \emph{arXiv:1606.06328 {[}Stat{]}}.
Retrieved from \url{http://arxiv.org/abs/1606.06328}

\hypertarget{ref-_st-dbscan:_2007}{}
Birant, D., \& Kut, A. (2007). ST-dbscan: An algorithm for clustering
spatial--temporal data. \emph{Data \& Knowledge Engineering},
\emph{60}(1), 208--221.

\hypertarget{ref-R-rgdal}{}
Bivand, R., Keitt, T., \& Rowlingson, B. (2017). \emph{Rgdal: Bindings
for the 'geospatial' data abstraction library}. Retrieved from
\url{https://CRAN.R-project.org/package=rgdal}

\hypertarget{ref-brunsdon_path_2007}{}
Brunsdon, C. (2007). Path estimation from GPS tracks. \emph{Proceedings
of the 9th International Conference on GeoComputation}. Retrieved from
\url{http://eprints.maynoothuniversity.ie/6148/}

\hypertarget{ref-chen_practical_2006}{}
Chen, M. Y., Sohn, T., Chmelev, D., Haehnel, D., Hightower, J., Hughes,
J., \ldots{} Varshavsky, A. (2006). Practical metropolitan-scale
positioning for GSM phones. In \emph{UbiComp 2006: Ubiquitous computing}
(pp. 225--242). Springer, Berlin, Heidelberg.
doi:\href{https://doi.org/10.1007/11853565_14}{10.1007/11853565\_14}

\hypertarget{ref-chen_state_2013}{}
Chen, Z., \& Brown, E. N. (2013). State space model.
\emph{Scholarpedia}, \emph{8}(3), 30868.
doi:\href{https://doi.org/10.4249/scholarpedia.30868}{10.4249/scholarpedia.30868}

\hypertarget{ref-R-leaflet}{}
Cheng, J., Karambelkar, B., \& Xie, Y. (n.d.). \emph{Leaflet: Create
interactive web maps with the javascript 'leaflet' library}. Retrieved
from \url{http://rstudio.github.io/leaflet/}

\hypertarget{ref-delclos-alio_keeping_2017}{}
Delclòs-Alió, X., Marquet, O., \& Miralles-Guasch, C. (2017). Keeping
track of time: A smartphone-based analysis of travel time perception in
a suburban environment. \emph{Travel Behaviour and Society},
\emph{9}(Supplement C), 1--9.
doi:\href{https://doi.org/10.1016/j.tbs.2017.07.001}{10.1016/j.tbs.2017.07.001}

\hypertarget{ref-duncan_portable_2013}{}
Duncan, S., Stewart, T. I., Oliver, M., Mavoa, S., MacRae, D., Badland,
H. M., \& Duncan, M. J. (2013). Portable global positioning system
receivers: Static validity and environmental conditions. \emph{American
Journal of Preventive Medicine}, \emph{44}(2), e19--29.
doi:\href{https://doi.org/10.1016/j.amepre.2012.10.013}{10.1016/j.amepre.2012.10.013}

\hypertarget{ref-commission_protecting_2017}{}
European Commission. (2017). \emph{Protecting your data: Your rights -
european commission}. Retrieved from
\url{http://ec.europa.eu/justice/data-protection/individuals/rights/index_en.htm}

\hypertarget{ref-feng_cutoff:_2014}{}
Feng, L., Nowak, G., O'Neill, T., \& Welsh, A. (2014). CUTOFF: A
spatio-temporal imputation method. \emph{Journal of Hydrology},
\emph{519}, 3591--3605.
doi:\href{https://doi.org/10.1016/j.jhydrol.2014.11.012}{10.1016/j.jhydrol.2014.11.012}

\hypertarget{ref-goldberg_international_2006}{}
Goldberg, L. R., Johnson, J. A., Eber, H. W., Hogan, R., Ashton, M. C.,
Cloninger, C. R., \& Gough, H. G. (2006). The international personality
item pool and the future of public-domain personality measures.
\emph{Journal of Research in Personality}, \emph{40}(1), 84--96.

\hypertarget{ref-goodchild_toward_2010}{}
Goodchild, M. F., \& Janelle, D. G. (2010). Toward critical spatial
thinking in the social sciences and humanities. \emph{GeoJournal},
\emph{75}(1), 3--13.
doi:\href{https://doi.org/10.1007/s10708-010-9340-3}{10.1007/s10708-010-9340-3}

\hypertarget{ref-grunerbl_smartphone-based_2015}{}
Grünerbl, A., Muaremi, A., Osmani, V., Bahle, G., Ohler, S., Tröster,
G., \ldots{} Lukowicz, P. (2015). Smartphone-based recognition of states
and state changes in bipolar disorder patients. \emph{IEEE Journal of
Biomedical and Health Informatics}, \emph{19}(1), 140--148.
doi:\href{https://doi.org/10.1109/JBHI.2014.2343154}{10.1109/JBHI.2014.2343154}

\hypertarget{ref-harari_using_2016}{}
Harari, G., Lane, N. D., Wang, R., Crosier, B. S., Campbell, A. T., \&
Gosling, S. D. (2016). Using smartphones to collect behavioral data in
psychological science: Opportunities, practical considerations, and
challenges. \emph{Perspectives on Psychological Science}, \emph{11}(6),
838--854.
doi:\href{https://doi.org/10.1177/1745691616650285}{10.1177/1745691616650285}

\hypertarget{ref-R-geosphere}{}
Hijmans, R. J. (2017a). \emph{Geosphere: Spherical trigonometry}.
Retrieved from \url{https://CRAN.R-project.org/package=geosphere}

\hypertarget{ref-R-raster}{}
Hijmans, R. J. (2017b). \emph{Raster: Geographic data analysis and
modeling}. Retrieved from
\url{https://CRAN.R-project.org/package=raster}

\hypertarget{ref-jankowska_framework_2015}{}
Jankowska, M. M., Schipperijn, J., \& Kerr, J. (2015). A framework for
using GPS data in physical activity and sedentary behavior studies.
\emph{Exercise and Sport Sciences Reviews}, \emph{43}(1), 48--56.
doi:\href{https://doi.org/10.1249/JES.0000000000000035}{10.1249/JES.0000000000000035}

\hypertarget{ref-lamarca_place_2005}{}
LaMarca, A., Chawathe, Y., Consolvo, S., Hightower, J., Smith, I.,
Scott, J., \ldots{} Schilit, B. (2005). Place lab: Device positioning
using radio beacons in the wild. In \emph{Pervasive computing} (pp.
116--133). Springer, Berlin, Heidelberg.
doi:\href{https://doi.org/10.1007/11428572_8}{10.1007/11428572\_8}

\hypertarget{ref-liao_learning_2007}{}
Liao, L., Patterson, D. J., Fox, D., \& Kautz, H. (2007). Learning and
inferring transportation routines. \emph{Artificial Intelligence},
\emph{171}(5), 311--331.
doi:\href{https://doi.org/10.1016/j.artint.2007.01.006}{10.1016/j.artint.2007.01.006}

\hypertarget{ref-london_encoding_2016}{}
London, I. (2016). \emph{Encoding cyclical continuous features - 24-hour
time. Ian london's blog}. Retrieved April 27, 2018, from
\url{//ianlondon.github.io/blog/encoding-cyclical-features-24hour-time/}

\hypertarget{ref-R-bindrcpp}{}
Müller, K. (2017). \emph{Bindrcpp: An 'rcpp' interface to active
bindings}. Retrieved from
\url{https://CRAN.R-project.org/package=bindrcpp}

\hypertarget{ref-R-RColorBrewer}{}
Neuwirth, E. (2014). \emph{RColorBrewer: ColorBrewer palettes}.
Retrieved from \url{https://CRAN.R-project.org/package=RColorBrewer}

\hypertarget{ref-palmius_detecting_2017}{}
Palmius, N., Tsanas, A., Saunders, K. E. A., Bilderbeck, A. C., Geddes,
J. R., Goodwin, G. M., \& Vos, M. D. (2017). Detecting bipolar
depression from geographic location data. \emph{IEEE Transactions on
Biomedical Engineering}, \emph{64}(8), 1761--1771.
doi:\href{https://doi.org/10.1109/TBME.2016.2611862}{10.1109/TBME.2016.2611862}

\hypertarget{ref-patterson_statespace_2008}{}
Patterson, T. A., Thomas, L., Wilcox, C., Ovaskainen, O., \&
Matthiopoulos, J. (2008). State--space models of individual animal
movement. \emph{Trends in Ecology \& Evolution}, \emph{23}(2), 87--94.
doi:\href{https://doi.org/10.1016/j.tree.2007.10.009}{10.1016/j.tree.2007.10.009}

\hypertarget{ref-R-sp}{}
Pebesma, E. J., \& Bivand, R. S. (2005). Classes and methods for spatial
data in R. \emph{R News}, \emph{5}(2), 9--13. Retrieved from
\url{https://CRAN.R-project.org/doc/Rnews/}

\hypertarget{ref-preisler_modeling_2004}{}
Preisler, H. K., Ager, A. A., Johnson, B. K., \& Kie, J. G. (2004).
Modeling animal movements using stochastic differential equations.
\emph{Environmetrics 15: P. 643-657}. Retrieved from
\url{https://www.fs.usda.gov/treesearch/pubs/33038}

\hypertarget{ref-R-base}{}
R Core Team. (2017). \emph{R: A language and environment for statistical
computing}. Vienna, Austria: R Foundation for Statistical Computing.
Retrieved from \url{https://www.R-project.org/}

\hypertarget{ref-sadilek_far_2016}{}
Sadilek, A., \& Krumm, J. (2016). Far out: Predicting long-term human
mobility. \emph{Microsoft Research}. Retrieved from
\url{https://www.microsoft.com/en-us/research/publication/far-predicting-long-term-human-mobility/}

\hypertarget{ref-saeb_mobile_2015}{}
Saeb, S., Zhang, M., Karr, C. J., Schueller, S. M., Corden, M. E.,
Kording, K. P., \& Mohr, D. C. (2015). Mobile phone sensor correlates of
depressive symptom severity in daily-life behavior: An exploratory
study. \emph{Journal of Medical Internet Research}, \emph{17}(7), e175.
doi:\href{https://doi.org/10.2196/jmir.4273}{10.2196/jmir.4273}

\hypertarget{ref-schipperijn_dynamic_2014}{}
Schipperijn, J., Kerr, J., Duncan, S., Madsen, T., Klinker, C. D., \&
Troelsen, J. (2014). Dynamic accuracy of GPS receivers for use in health
research: A novel method to assess GPS accuracy in real-world settings.
\emph{Frontiers in Public Health}, \emph{2}, 21.
doi:\href{https://doi.org/10.3389/fpubh.2014.00021}{10.3389/fpubh.2014.00021}

\hypertarget{ref-simmons_false-positive_2011}{}
Simmons, J. P., Nelson, L. D., \& Simonsohn, U. (2011). False-positive
psychology: Undisclosed flexibility in data collection and analysis
allows presenting anything as significant. \emph{Psychological Science},
\emph{22}(11), 1359--1366.
doi:\href{https://doi.org/10.1177/0956797611417632}{10.1177/0956797611417632}

\hypertarget{ref-sobrado_my_2018}{}
Sobrado, B. (2018). \emph{My thesis on missing data in GPS
measurements.} Retrieved from \url{https://github.com/sobradob/thesis}

\hypertarget{ref-tatsiramos_geographic_2009}{}
Tatsiramos, K. (2009). Geographic labour mobility and unemployment
insurance in europe. \emph{Journal of Population Economics},
\emph{22}(2), 267--283.
doi:\href{https://doi.org/10.1007/s00148-008-0194-7}{10.1007/s00148-008-0194-7}

\hypertarget{ref-R-padr}{}
Thoen, E. (2017). \emph{Padr: Quickly get datetime data ready for
analysis}. Retrieved from \url{https://CRAN.R-project.org/package=padr}

\hypertarget{ref-location_history_timeline_2017}{}
\emph{Timeline}. (2017). Retrieved from
\url{https://www.google.com/maps/timeline?pb}

\hypertarget{ref-R-tibbletime}{}
Vaughan, D., \& Dancho, M. (2018). \emph{Tibbletime: Time aware
tibbles}. Retrieved from
\url{https://CRAN.R-project.org/package=tibbletime}

\hypertarget{ref-villanueva_association_2013}{}
Villanueva, C., \& Aggarwal, B. (2013). The association between
neighborhood socioeconomic status and clinical outcomes among patients 1
year after hospitalization for cardiovascular disease. \emph{Journal of
Community Health}, \emph{38}(4), 690--697.
doi:\href{https://doi.org/10.1007/s10900-013-9666-0}{10.1007/s10900-013-9666-0}

\hypertarget{ref-wang_smartgpa:_2015}{}
Wang, R., Harari, G., Hao, P., Zhou, X., \& Campbell, A. T. (2015).
SmartGPA: How smartphones can assess and predict academic performance of
college students. In \emph{Proceedings of the 2015 ACM international
joint conference on pervasive and ubiquitous computing} (pp. 295--306).
New York, NY, USA: ACM.
doi:\href{https://doi.org/10.1145/2750858.2804251}{10.1145/2750858.2804251}

\hypertarget{ref-R-ggplot2}{}
Wickham, H. (2009). \emph{Ggplot2: Elegant graphics for data analysis}.
Springer-Verlag New York. Retrieved from \url{http://ggplot2.org}

\hypertarget{ref-R-scales}{}
Wickham, H. (2017). \emph{Scales: Scale functions for visualization}.
Retrieved from \url{https://CRAN.R-project.org/package=scales}

\hypertarget{ref-R-tidyr}{}
Wickham, H., \& Henry, L. (2018). \emph{Tidyr: Easily tidy data with
'spread()' and 'gather()' functions}. Retrieved from
\url{https://CRAN.R-project.org/package=tidyr}

\hypertarget{ref-R-dbplyr}{}
Wickham, H., \& Ruiz, E. (2018). \emph{Dbplyr: A 'dplyr' back end for
databases}. Retrieved from
\url{https://CRAN.R-project.org/package=dbplyr}

\hypertarget{ref-R-dplyr}{}
Wickham, H., Francois, R., Henry, L., \& Müller, K. (2017). \emph{Dplyr:
A grammar of data manipulation}. Retrieved from
\url{https://CRAN.R-project.org/package=dplyr}

\hypertarget{ref-R-readr}{}
Wickham, H., Hester, J., \& Francois, R. (2017). \emph{Readr: Read
rectangular text data}. Retrieved from
\url{https://CRAN.R-project.org/package=readr}

\hypertarget{ref-wolf_impact_2003}{}
Wolf, J., Oliveira, M., \& Thompson, M. (2003). Impact of underreporting
on mileage and travel time estimates: Results from global positioning
system-enhanced household travel survey. \emph{Transportation Research
Record: Journal of the Transportation Research Board}, \emph{1854},
189--198. doi:\href{https://doi.org/10.3141/1854-21}{10.3141/1854-21}

\hypertarget{ref-wu_spatial-temporal-semantic_2017}{}
Wu, F., Fu, K., Wang, Y., Xiao, Z., \& Fu, X. (2017). A
spatial-temporal-semantic neural network algorithm for location
prediction on moving objects. \emph{Algorithms}, \emph{10}(2), 37.
doi:\href{https://doi.org/10.3390/a10020037}{10.3390/a10020037}

\hypertarget{ref-R-knitr}{}
Xie, Y. (2015). \emph{Dynamic documents with R and knitr} (2nd ed.).
Boca Raton, Florida: Chapman; Hall/CRC. Retrieved from
\url{https://yihui.name/knitr/}

\hypertarget{ref-zenk_how_2009}{}
Zenk, S. N., Schulz, A. J., \& Odoms-Young, A. (2009). How neighborhood
environments contribute to obesity. \emph{The American Journal of
Nursing}, \emph{109}(7), 61--64.
doi:\href{https://doi.org/10.1097/01.NAJ.0000357175.86507.c8}{10.1097/01.NAJ.0000357175.86507.c8}

\hypertarget{ref-zhang_application_2017}{}
Zhang, Z., Yang, X., Li, H., Li, W., Yan, H., \& Shi, F. (2017).
Application of a novel hybrid method for spatiotemporal data imputation:
A case study of the minqin county groundwater level. \emph{Journal of
Hydrology}, \emph{553}(Supplement C), 384--397.
doi:\href{https://doi.org/10.1016/j.jhydrol.2017.07.053}{10.1016/j.jhydrol.2017.07.053}

\hypertarget{ref-R-kableExtra}{}
Zhu, H. (2018). \emph{KableExtra: Construct complex table with 'kable'
and pipe syntax}. Retrieved from
\url{https://CRAN.R-project.org/package=kableExtra}

\endgroup

\end{document}